---
title: "Assignment 2 Markdown"
author: "Mark Lewis"
date: "March 9, 2019"
output: pdf_document
---

```{r}
#copy above to insert a new block

library(dplyr)
library(reshape2)
library(ggplot2)
library(Hmisc)
library(corrplot)
library(mice)
library(VIM)
library(pROC)
library(caret)
library(sqldf)
library(readxl)
library(ggplot2)

library(dbscan)
library(fpc)
library(factoextra)
library(NbClust)
library(vegan)
```


#Import the raw data set

```{r}
data <- read.csv("OnlineRetail.csv") 
summary(data)
```
#Create a new collumn containing total amount of money spent

```{r}
data$totalSpent <- data$UnitPrice * data$Quantity
```


#Convert the dates from strings to the appropriate format so they can be interpreted as such
```{r}
dates <- as.character(data$InvoiceDate)
datesX <- strsplit(dates, " ")
datesX <- matrix(unlist(datesX), ncol=2, byrow=TRUE)
datesY <- strsplit(datesX[,1], "/")
datesY <- matrix(unlist(datesY), ncol=3, byrow=TRUE)
datesZ <- strsplit(datesX[,2], ":")
datesZ <- matrix(unlist(datesZ), ncol=2, byrow=TRUE)
```

#Create new columns for Year, Month, Day, Hour and Minute
```{r}
data$month <- datesY[,1]
data$day <- datesY[,2]
data$year <- datesY[,3]
data$hour <- datesZ[,1]
data$minute <- datesZ[,2]


data$month <- as.numeric(as.character(data$month))
data$day <- as.numeric(as.character(data$day))
data$year <- as.numeric(as.character(data$year))
data$hour <- as.numeric(as.character(data$hour))
data$minute <- as.numeric(as.character(data$minute))

```

#DATA EXPLORATION


#ORDERS

#How many unique orders are there? 
```{r }
str(data$InvoiceNo)
summary(data$InvoiceNo)
```

#What is the average number of products in an Order?
```{r }
#Dont run this takes to long 
#OrderTable <- levels(data$InvoiceNo)
#OrderTable <- cbind(OrderTable)
#OrderTable <- as.data.frame(as.table(OrderTable))
#OrderTable$Var1 <- NULL
#OrderTable$Var2 <- NULL
#OrderTable[,"AverageNumofProd"] <- NA

#for( i in rownames(OrderTable) ){
#  for( j in rownames(data)){
#    if(as.character(OrderTable[i,"Freq"])==as.character(data[j,"InvoiceNo"])){
#      if(is.na(OrderTable[i,"AverageNumofProd"])){
#        OrderTable[i,"AverageNumofProd"] <- 1
#      }
#      else{
#        OrderTable[i,"AverageNumofProd"] <- OrderTable[i,"AverageNumofProd"]+1
#      }
#    }
#  }
#}
```


#What are the Minimum and Maximum Amounts spend on Orders?

```{r }
min(data$totalSpent)
max(data$totalSpent)
```

#What is the Average Total Amount and Median Total Amount spent on each Order?
```{r }
summary(data$totalSpent)
```

#PRODUCTS

#How many unique Products are there?
```{r }
str(data$StockCode)
summary(data$StockCode)
str(data$Description)
```

#What are the Minimum and Maximum Prices of Products?
```{r }
summary(data$UnitPrice)
```
#What are the Average and Median prices of these Products? 
```{r }
mean(data$UnitPrice)
median(data$UnitPrice)
```


#CUSTOMERS

#How many unique customers are there?
```{r }
data$CustomerID <- as.factor(data$CustomerID)
str(data$CustomerID)
```

#Where are the customers from?
```{r }
summary(data$Country)
```
#What is the Minimum amount, Maximum Amount, Average and Median amounts spent by the Customer?
```{r }
#Dont run this takes to long 
# Customers <- levels(data$CustomerID)
# Customers <- cbind(Customers)
# Customers <- as.data.frame(as.table(Customers))
# Customers$Var1 <- NULL
# Customers$Var2 <- NULL
# Customers[,"AmountSpent"] <- NA
# for( i in rownames(Customers) ){
#   for( j in rownames(data)){
#     if(as.character(Customers[i,"Freq"])==as.character(data[j,"CustomerID"])){
#       if(is.na(Customers[i,"AmountSPent"])){
#         Customers[i,"AmountSpent"] <- data[j,"totalSpent"]
#       }
#       else{
#         Customers[i,"AmountSpent"] <- (Customers[i,"totalSpent"]+data[j,"totalSpent"])/2
#       }
#     }
#   }
# }
# summary(Customers$AmountSpent)
```

#WILL CLEAN UP ROWS 120 - 184
#creating new collumns for new transactions
```{r }
data$Cancelled<-ifelse(substring(data$InvoiceNo,0,1)=="C",1,0)
data$Amazon<-ifelse(data$StockCode =="AMAZONFEE",1,0)
data$BANK<-ifelse(data$StockCode =="BANK CHARGES",1,0)
data$Carriage<-ifelse(data$StockCode =="C2",1,0)
data$Commision<-ifelse(data$StockCode =="CRUK",1,0)
data$Discount<-ifelse(data$StockCode =="D",1,0)
data$DOT<-ifelse(data$StockCode =="DOT",1,0)
data$Manual<-ifelse(data$StockCode =="M",1,0)
data$post<-ifelse(data$StockCode =="POST",1,0)
data$Samples<-ifelse(data$StockCode =="SAMPLES",1,0)
```
#general visualizations of each feature
#DATA EXPLORATION HERE 
```{r }
summary(data$Description)
```
#you can see the the most popular items are listed first
```{r }
str(data$Description)
summary(data$Quantity)
table(data$Quantity)
summary(data$UnitPrice)
summary(data$Country)
summary(data$totalSpent)
table(data$month,data$day)
table(data$year,data$month)
```
#example of a table you can print to see which products are popular depending on the month
```{r }
table(data$Description,data$month)
```
#a few more examples
```{r }
table(data$Country,data$hour)
```
#the lie below allows for the full set to be printed out
```{r }
#options(max.print=999999)
```
#here you can get an image of the most popular products

#visualization of unique transactions
#they take up only a small portion of transactions in the dataset
```{r }
table(data$Cancelled)
table(data$Amazon)
table(data$BANK)
table(data$Carriage)
table(data$Commision)
table(data$Commision)
table(data$Discount)
table(data$DOT)
table(data$Manual)
table(data$post)
table(data$Samples)

summary(data$Quantity)
summary(data$UnitPrice)
```

#graph

#view number of cancelled transactions
```{r }
table(data$Cancelled)
```
#missing data
```{r }
is.na(data) = data==""
sum(is.na(data))
```
#sum of rows missing from data
```{r }
mean(is.na(data))
```



#DATA PREPARATION

#1. Remove Cancelled Transactions, Invalid Descriptions, and missing values
```{r }
data<-data[!(data$Amazon=="1" & data$BANK=="1" &data$Cancelled=="1" &data$Carriage=="1" &data$Commision=="1" &data$Discount=="1" &data$DOT=="1" &data$Manual=="1" &data$post=="1" &data$Samples=="1"),]
data = na.omit(data)
```




#adjust dataset so that it can be interperated by unsupervised learning algorithms
```{r }
data$InvoiceNo <- NULL
data$StockCode<-NULL
data$Description<-NULL
data$InvoiceDate<-NULL
data$Country<-NULL
data$CustomerID<-NULL
data$year<-data$year-2010
data$totalSpent<-NULL
data$Cancelled <- NULL
data$Amazon <- NULL
data$BANK<-NULL
data$Carriage<-NULL
data$Commision<-NULL
data$Discount<-NULL
data$DOT<-NULL
data$Manual<-NULL
data$post<-NULL
data$Samples<-NULL
```
#deal with outliers
```{r }
data$UnitPrice<-ifelse(data$UnitPrice < 0,-50,data$UnitPrice)
data$Quantity<-ifelse(data$Quantity < 0,-50,data$Quantity)
```

#RUNNING THE MODEL

#Determine number of clusters
```{r }
model <- cascadeKM(data, 1, 10, iter = 10)#takes to long to run with more iters dont bother
model$results[2,]
```
#Plot this following line last takes to long to generate plot
#Problems seem to arise since the data set is so big R studio doesnt seem to handle all the memory
```{r }
#plot(model, sortg = TRUE)
```
#Recommends which number of clusters to use, outputs 3
```{r }
which.max(model$results[2,])
```
#We're going with 2 clusters
```{r }
kresult <- kmeans(data,3)#3 if missing data is omiited #2if data wasnt omitted
kresult$size
```
#see the visualization of the spread of clusters
#Add clusters as a collumn corresponding to each transaction

#other algorithms doesnt work because the data set it to big quotes the vector size is too big
```{r }
#hclustresult <- hclust(dist(data))
#dbscanresult<- dbscan(data, .15)
```

